Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)

### 2025-07-11 resubmission/paper-blinded-resubmit.qmd
Update carpeta resubmission y varios (archivos no renderizados)

---
title: "Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)"
format:
  html: 
    code-fold: true
    html-math-method: katex
    number-sections: true
  docx:
    number-sections: true
  pdf: 
    geometry: margin=2cm
    template-partials: 
      - title.tex
    keep-tex: true
    include-in-header:
      text: |
        \usepackage[noblocks]{authblk}
        \renewcommand*{\Authsep}{, }
        \renewcommand*{\Authand}{, }
        \renewcommand*{\Authands}{, }
        \renewcommand\Affilfont{\small}
    number-sections: true
editor: source
bibliography: "input/bib/merit-edjust.bib"
# csl: "input/bib/apa6.csl"
csl: "input/bib/harvard.csl"
author:
  - name: Juan Carlos Castillo
    corresponding: true
    orcid: 0000-0003-1265-7854
    email: juancastillov@uchile.cl
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 3
  - name: Andreas Laffert
    orcid: 0000-0002-9008-2454
    email: alaffertt@estudiante.uc.cl 
    affiliations:
      - ref: 4
  - name: Kevin Carrasco
    orcid: 0000-0002-9348-8361
    email: kevin.carrasco@ug.uchile.cl
    affiliations:
      - ref: 2
  - name: Julio Iturra-Sanhueza
    orcid: 0000-0003-1993-3479
    email: julioiturrasanhueza@gmail.com 
    affiliations:
      - ref: 5     
affiliations: 
  - id: 1
    name: Departamento de Sociología, Universidad de Chile
  - id: 2
    name: Centro de estudios del conflicto y cohesión social (COES)
  - id: 3
    name: Núcleo milenio de desigualdades y oportunidades digitales (NUDOS)
  - id: 4
    name: Instituto de Sociología, Pontificia Universidad Católica de Chile
  - id: 5
    name: International Graduate School of Social Sciencies (BIGSSS), University of Bremen, Germany
citeproc: true
abstract: | 
  This study investigates the relationship between perceptions of economic inequality, meritocratic perceptions, and preferences for market justice in Chile between 2016 and 2023. Using six waves of panel data from the Chilean Longitudinal Social Survey - ELSOC ($N_{observations}$ = 8,643; $N_{individuals}$ = 1,687), the analysis examines how subjective assessments of inequality shape attitudes toward the role of merit in access to key social services such as healthcare, education, and pensions. Results indicate that rising perceptions of inequality are associated with lower support for market justice; however, individuals with strong meritocratic convictions are more likely to legitimize existing disparities. The study also considers the influence of major social movements during this period, which appear to have reshaped public discourse and perceptions of fairness. These findings contribute to a deeper understanding of how beliefs about justice and equity evolve in contexts marked by persistent inequality and entrenched market-oriented frameworks.
  \newline
  **Keywords**: Economic inequality, meritocracy, market justice, Chile, public preferences, inequality perception
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
colorlinks: true
fig-height: 4
fig-width: 7.5
---

>>>>

---
title: "Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)"
format:
  html: 
    code-fold: true
    html-math-method: katex
    number-sections: true
  docx:
    number-sections: true
  pdf: 
    geometry: margin=2cm
    template-partials: 
      - title.tex
    keep-tex: true
    include-in-header:
      text: |
        \usepackage[noblocks]{authblk}
        \renewcommand*{\Authsep}{, }
        \renewcommand*{\Authand}{, }
        \renewcommand*{\Authands}{, }
        \renewcommand\Affilfont{\small}
    number-sections: true
editor: source
bibliography: "../input/bib/merit-edjust.bib"
# csl: "input/bib/apa6.csl"
csl: "../input/bib/harvard.csl"
# author:
#   - name: Juan Carlos Castillo
#     corresponding: true
#     orcid: 0000-0003-1265-7854
#     email: juancastillov@uchile.cl
#     affiliations:
#       - ref: 1
#       - ref: 2
#       - ref: 3
#   - name: Andreas Laffert
#     orcid: 0000-0002-9008-2454
#     email: alaffertt@estudiante.uc.cl 
#     affiliations:
#       - ref: 4
#   - name: Kevin Carrasco
#     orcid: 0000-0002-9348-8361
#     email: kevin.carrasco@ug.uchile.cl
#     affiliations:
#       - ref: 2
#   - name: Julio Iturra-Sanhueza
#     orcid: 0000-0003-1993-3479
#     email: julioiturrasanhueza@gmail.com 
#     affiliations:
#       - ref: 5     
# affiliations: 
#   - id: 1
#     name: Departamento de Sociología, Universidad de Chile
#   - id: 2
#     name: Centro de estudios del conflicto y cohesión social (COES)
#   - id: 3
#     name: Núcleo milenio de desigualdades y oportunidades digitales (NUDOS)
#   - id: 4
#     name: Instituto de Sociología, Pontificia Universidad Católica de Chile
#   - id: 5
#     name: International Graduate School of Social Sciencies (BIGSSS), University of Bremen, Germany
citeproc: true
abstract: | 
 This study investigates the relationship between perceptions of economic inequality, meritocratic perceptions, and preferences for market justice in Chile between 2016 and 2023. Using six waves of panel data from the Chilean Longitudinal Social Survey - ELSOC ($N_{observations}$ = 8,643; $N_{individuals}$ = 1,687), the analysis examines how subjective assessments of inequality shape attitudes toward the role of merit in access to key social services such as healthcare, education, and pensions. Results indicate that rising perceptions of inequality are associated with lower support for market justice; however, individuals with strong meritocratic convictions are more likely to legitimize existing disparities. The study also considers the influence of major social movements during this period, which appear to have reshaped public discourse and perceptions of fairness. These findings contribute to a deeper understanding of how beliefs about justice and equity evolve in contexts marked by persistent inequality and entrenched market-oriented frameworks.
  \newline
  **Keywords**: Economic inequality, meritocracy, market justice, Chile, public preferences, inequality perception
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
colorlinks: true
fig-height: 4
fig-width: 7.5
---

### 2025-07-11 resubmission/paper-resubmit.qmd
Update carpeta resubmission y varios (archivos no renderizados)

---
title: "Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)"
format:
  html: 
    code-fold: true
    html-math-method: katex
    number-sections: true
  docx:
    number-sections: true
  pdf: 
    geometry: margin=2cm
    template-partials: 
      - title.tex
    keep-tex: true
    include-in-header:
      text: |
        \usepackage[noblocks]{authblk}
        \renewcommand*{\Authsep}{, }
        \renewcommand*{\Authand}{, }
        \renewcommand*{\Authands}{, }
        \renewcommand\Affilfont{\small}
    number-sections: true
editor: source
bibliography: "../input/bib/merit-edjust.bib"
# csl: "input/bib/apa6.csl"
csl: "../input/bib/harvard.csl"
# author:
#   - name: Juan Carlos Castillo
#     corresponding: true
#     orcid: 0000-0003-1265-7854
#     email: juancastillov@uchile.cl
#     affiliations:
#       - ref: 1
#       - ref: 2
#       - ref: 3
#   - name: Andreas Laffert
#     orcid: 0000-0002-9008-2454
#     email: alaffertt@estudiante.uc.cl 
#     affiliations:
#       - ref: 4
#   - name: Kevin Carrasco
#     orcid: 0000-0002-9348-8361
#     email: kevin.carrasco@ug.uchile.cl
#     affiliations:
#       - ref: 2
#   - name: Julio Iturra-Sanhueza
#     orcid: 0000-0003-1993-3479
#     email: julioiturrasanhueza@gmail.com 
#     affiliations:
#       - ref: 5     
# affiliations: 
#   - id: 1
#     name: Departamento de Sociología, Universidad de Chile
#   - id: 2
#     name: Centro de estudios del conflicto y cohesión social (COES)
#   - id: 3
#     name: Núcleo milenio de desigualdades y oportunidades digitales (NUDOS)
#   - id: 4
#     name: Instituto de Sociología, Pontificia Universidad Católica de Chile
#   - id: 5
#     name: International Graduate School of Social Sciencies (BIGSSS), University of Bremen, Germany
citeproc: true
abstract: | 
 This study investigates the relationship between perceptions of economic inequality, meritocratic perceptions, and preferences for market justice in Chile between 2016 and 2023. Using six waves of panel data from the Chilean Longitudinal Social Survey - ELSOC ($N_{observations}$ = 8,643; $N_{individuals}$ = 1,687), the analysis examines how subjective assessments of inequality shape attitudes toward the role of merit in access to key social services such as healthcare, education, and pensions. Results indicate that rising perceptions of inequality are associated with lower support for market justice; however, individuals with strong meritocratic convictions are more likely to legitimize existing disparities. The study also considers the influence of major social movements during this period, which appear to have reshaped public discourse and perceptions of fairness. These findings contribute to a deeper understanding of how beliefs about justice and equity evolve in contexts marked by persistent inequality and entrenched market-oriented frameworks.
  \newline
  **Keywords**: Economic inequality, meritocracy, market justice, Chile, public preferences, inequality perception
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
colorlinks: true
fig-height: 4
fig-width: 7.5
---

>>>>

---
title: "Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)"
format:
  html: 
    code-fold: true
    html-math-method: katex
    number-sections: true
  docx:
    number-sections: true
  pdf: 
    geometry: margin=2cm
    template-partials: 
      - title.tex
    keep-tex: true
    include-in-header:
      text: |
        \usepackage[noblocks]{authblk}
        \renewcommand*{\Authsep}{, }
        \renewcommand*{\Authand}{, }
        \renewcommand*{\Authands}{, }
        \renewcommand\Affilfont{\small}
    number-sections: true
editor: source
bibliography: "../input/bib/merit-edjust.bib"
# csl: "input/bib/apa6.csl"
csl: "../input/bib/harvard.csl"
author:
  - name: Juan Carlos Castillo
    corresponding: true
    orcid: 0000-0003-1265-7854
    email: juancastillov@uchile.cl
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 3
  - name: Andreas Laffert
    orcid: 0000-0002-9008-2454
    email: alaffertt@estudiante.uc.cl 
    affiliations:
      - ref: 4
  - name: Kevin Carrasco
    orcid: 0000-0002-9348-8361
    email: kevin.carrasco@ug.uchile.cl
    affiliations:
      - ref: 2
  - name: Julio Iturra-Sanhueza
    orcid: 0000-0003-1993-3479
    email: julioiturrasanhueza@gmail.com 
    affiliations:
      - ref: 5     
affiliations: 
  - id: 1
    name: Departamento de Sociología, Universidad de Chile
  - id: 2
    name: Centro de estudios del conflicto y cohesión social (COES)
  - id: 3
    name: Núcleo milenio de desigualdades y oportunidades digitales (NUDOS)
  - id: 4
    name: Instituto de Sociología, Pontificia Universidad Católica de Chile
  - id: 5
    name: International Graduate School of Social Sciencies (BIGSSS), University of Bremen, Germany
citeproc: true
abstract: | 
  This study investigates the relationship between perceptions of economic inequality, meritocratic perceptions, and preferences for market justice in Chile between 2016 and 2023. Using six waves of panel data from the Chilean Longitudinal Social Survey - ELSOC ($N_{observations}$ = 8,643; $N_{individuals}$ = 1,687), the analysis examines how subjective assessments of inequality shape attitudes toward the role of merit in access to key social services such as healthcare, education, and pensions. Results indicate that rising perceptions of inequality are associated with lower support for market justice; however, individuals with strong meritocratic convictions are more likely to legitimize existing disparities. The study also considers the influence of major social movements during this period, which appear to have reshaped public discourse and perceptions of fairness. These findings contribute to a deeper understanding of how beliefs about justice and equity evolve in contexts marked by persistent inequality and entrenched market-oriented frameworks.
  \newline
  **Keywords**: Economic inequality, meritocracy, market justice, Chile, public preferences, inequality perception
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
colorlinks: true
fig-height: 4
fig-width: 7.5
---
